---
title: "Project 2"
subtitle: "Proposal"
authors: ""
format: html
editor: visual
---

```{r load-packages}
#| label: load-pkgs
#| message: false
#| warning: false

library(tidyverse)
```

## Datasets, Web Sources, and Packages

```{r load-data}
#| label: load-data
#| message: false

## read data 
```

## Why Did We Choose this Data?

We chose to use `rvest`, `textclassifier`, and `rcrawler`(with the support of the `tidyverse` R package) to retrieve, clean, and interpret text of interest publicly available online . These packages in particular are good for keeping our data wrangling lean, which is preferred since our data of interest is disparate in nature. We will be pulling much of our data from niche.com, since many student reviews online reside on this website. Niche's formatting of their reviews will be useful as well, since they come with crucial identifiers such as year a student is at the time of posting(freshman, junior, alum, etc), and the time they posted. We may also use datasets such as the College Scorecard offered by the U.S. Department of Education, to supplement our analysis. Additionally, we will be using the NRC Emotion Lexicon, and Bing Lexicon to support our sentiment analysis. And finally, we will use the `Shiny` R package to visualize our results in an interactive dashboard.

# Questions and Analysis Plan

## How do students really feel about their respective colleges?

College admissions in the U.S. are becoming increasingly difficult to navigate; however, applications are on the rise. Students are left to embark on their own college searches, with varying levels of guidance and support. Despite the abundance of *college search* tools available, many rely on aging metrics and flimsy statistics(such as standardized testing scores, and college "prestige") to aid an applicants' derision-making process. However, internal data(for Reed), tells us that two of the most impactful factors in an applicant's cycle, is speaking to current students of an institution, and physically visiting campus. Both of which comprise what we deem, a college's *student sentiment score*. Getting "the vibe" of a college be challenging, as most students aren't able to visit campuses in-person before making a decision, and even if they could, access to honest, holistic, and representative student perspective is often limited.

By retrieving students opinions of their institution from online and predicatively interpreting that data, we aim to quantify a college's "vibe", and compile a set of sentiment scores that could help applicants better gauge their *fit* for a particulate institution. **Reed College will be the focus of our project, but we hope to expand to a broader array of schools as time allows.** We will be utilizing web crawling/scraping algorithms to pull whatever Reed students are saying about Reed online, within the scope of our analysis. Once tidied, we will use a predictive "text classifier" efficiently categorize student opinions for both general, and focused sentiment analysis(what students are saying about [Reed]{.underline} vs. what students are saying about [professor]{.underline} at Reed). And to conclude, we will visualize these scores on an interactive dashboard, where users could(as time allows) view student sentiment for several institutions over time.

**Our minimum viable product(MVP) for this project, will be a dashboard that displays a general sentiment score for a small set of institutions, over time.** Below is our weekly action plan:
