---
title: "CollegeScorecard Data Wrangling"
output: pdf_document
date: '2024-04-12'
---

```{r}
library(readr)
library(dplyr)
library(tidyverse)

setwd("/Users/ashwindev/Desktop/DataScience/project-02-rosa1/Processing/Wrangling")
```


```{r}
Most_Recent_Cohorts_Institution <- read_csv("Most-Recent-Cohorts-Institution.csv")
REED_TIDY <- read_csv("REED_TIDY.csv")
```

```{r}
head(Most_Recent_Cohorts_Institution)
tail(Most_Recent_Cohorts_Institution)
```


Important variables: 


ADM_RATE  
UGDS (Enrollment of undergraduate certificate/degree-seeking students)
UGDS_MEN
UGDS_WOMEN    
GRADS (# of graduate studnets)
CCSIZSET (Carnegie Classification -- size and setting)



UGDS_WHITE
UGDS_BLACK
UGDS_HISP
UGDS_ASIAN
UGDS_AIAN
UGDS_NHPI
UGDS_2MOR
UGDS_NRA
UGDS_UNKN
UGDS_WHITENH
UGDS_BLACKNH
UGDS_API
UGDS_AIANOLD
UGDS_HISPOLD
UG_NRA
UG_UNKN
UG_WHITENH
UG_BLACKNH
UG_API
UG_AIANOLD
UG_HISPOLD


TUITIONFEE_IN (in-state tuit)
TUITIONFEE_OUT (out of state tuit)
PCTPELL (% students receiving pell)
PCTFLOAN (% students receiving a student federal loan)
GRAD_DEBT_MDN (The median debt for students who have completed)
FAMINC (avg family income)
MD_FAMINC (median family income)

```{r}
college <- Most_Recent_Cohorts_Institution %>%
  select(
    ADM_RATE,
    UGDS,
    UGDS_MEN,
    UGDS_WOMEN,
    GRADS,
    CCSIZSET,
    UGDS_WHITE,
    UGDS_BLACK,
    UGDS_HISP,
    UGDS_ASIAN,
    UGDS_AIAN,
    UGDS_NHPI,
    UGDS_2MOR,
    UGDS_NRA,
    UGDS_UNKN,
    UGDS_WHITENH,
    UGDS_BLACKNH,
    UGDS_API,
    UGDS_AIANOLD,
    UGDS_HISPOLD,
    UG_NRA,
    UG_UNKN,
    UG_WHITENH,
    UG_BLACKNH,
    UG_API,
    UG_AIANOLD,
    UG_HISPOLD,
    TUITIONFEE_IN,
    TUITIONFEE_OUT,
    PCTPELL,
    PCTFLOAN,
    GRAD_DEBT_MDN,
    FAMINC,
    MD_FAMINC
  )
```

EDA

```{r}
college %>% 
  summarise(
    Average_Tuition_In = mean(TUITIONFEE_IN, na.rm = TRUE),
    Average_Tuition_Out = mean(TUITIONFEE_OUT, na.rm = TRUE),
    Median_Family_Income = median(MD_FAMINC, na.rm = TRUE),
    Average_Pct_Pell = mean(PCTPELL, na.rm = TRUE)
  )
```

```{r}
# Ensure TUITIONFEE_IN is numeric
college$TUITIONFEE_IN <- as.numeric(as.character(college$TUITIONFEE_IN))

# Histogram for In-State Tuition Fees
ggplot(college, aes(x = TUITIONFEE_IN)) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7)

# Ensure MD_FAMINC is numeric
college$MD_FAMINC <- as.numeric(as.character(college$MD_FAMINC))

# Histogram for Median Family Income
ggplot(college, aes(x = MD_FAMINC)) +
  geom_histogram(bins = 30, fill = "green", alpha = 0.7) +
  labs(title = "Distribution of Median Family Income", x = "Median Family Income", y = "Frequency")
```


```{r}
library(dplyr)
library(ggplot2)

# Convert factors to numeric if necessary
college <- college %>%
  mutate(
    UGDS_MEN = as.numeric(as.character(UGDS_MEN)),
    UGDS_WOMEN = as.numeric(as.character(UGDS_WOMEN))
  )

# Handle missing values if necessary, for example, replacing NA with 0
# This step depends on the context and how you want to treat missing values
college <- college %>%
  mutate(
    UGDS_MEN = ifelse(is.na(UGDS_MEN), 0, UGDS_MEN),
    UGDS_WOMEN = ifelse(is.na(UGDS_WOMEN), 0, UGDS_WOMEN)
  )

# Create Total_UGDS column
college <- college %>%
  mutate(Total_UGDS = UGDS_MEN + UGDS_WOMEN)

# Create the plot
college %>%
  ggplot(aes(x = factor(1), y = UGDS_MEN / Total_UGDS, fill = "Men")) +
  geom_bar(stat = "identity") +
  geom_bar(aes(y = UGDS_WOMEN / Total_UGDS, fill = "Women"), stat = "identity") +
  scale_fill_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(x = "", y = "Percentage", title = "Gender Distribution in Undergraduate Enrollment") +
  coord_polar(theta = "y")
```

```{r}
library(dplyr)

# Convert necessary columns to numeric
college <- college %>%
  mutate(
    TUITIONFEE_IN = as.numeric(as.character(TUITIONFEE_IN)),
    TUITIONFEE_OUT = as.numeric(as.character(TUITIONFEE_OUT)),
    MD_FAMINC = as.numeric(as.character(MD_FAMINC)),
    PCTPELL = as.numeric(as.character(PCTPELL)),
    PCTFLOAN = as.numeric(as.character(PCTFLOAN))
  )

# Selecting numeric variables
numeric_vars <- college %>% 
  select(TUITIONFEE_IN, TUITIONFEE_OUT, MD_FAMINC, PCTPELL, PCTFLOAN)

# Compute the correlation matrix with omitted NA values
correlation_matrix <- cor(na.omit(numeric_vars))
print(correlation_matrix)
```
