---
title: "CollegeScorecard Data Wrangling"
output: pdf_document
date: '2024-04-12'
---

```{r}
library(readr)
library(dplyr)
library(tidyverse)

setwd("/Users/ashwindev/Desktop/DataScience/project-02-rosa1/Processing/Wrangling")
```


```{r}
Most_Recent_Cohorts_Institution <- read_csv("Most-Recent-Cohorts-Institution.csv")
```

```{r}
head(Most_Recent_Cohorts_Institution)
tail(Most_Recent_Cohorts_Institution)
```

API key:

```{r}
library(rscorecard)

sc_key('hriAjENzUTJKzKt0zYxBvA6XfrWdHJwDum48RuoS')
```

```{r}
sc_dict(filter_vars = TRUE)
```

```{r}
data <- sc_init() %>%
  sc_filter(preddeg == 3, control == 1) %>%  # Filter for predominantly bachelor’s degree institutions and public control
  sc_select(UNITID, INSTNM, ADM_RATE, TUITIONFEE_IN, MEDIAN_HH_INC) %>%  # Selecting specific variables
  sc_year(2020) %>%
  sc_get()
```


```{r}
data <- sc_init() %>%
  sc_select(UNITID, INSTNM) %>%  # Just select institution ID and name to test
  sc_year(2020) %>%
  sc_get()

# Check what the data looks like
head(data)
```


Important variables: 


ADM_RATE  
UGDS (Enrollment of undergraduate certificate/degree-seeking students)
UGDS_MEN
UGDS_WOMEN    
GRADS (# of graduate studnets)
CCSIZSET (Carnegie Classification -- size and setting)



UGDS_WHITE
UGDS_BLACK
UGDS_HISP
UGDS_ASIAN
UGDS_AIAN
UGDS_NHPI
UGDS_2MOR
UGDS_NRA
UGDS_UNKN
UGDS_WHITENH
UGDS_BLACKNH
UGDS_API
UGDS_AIANOLD
UGDS_HISPOLD
UG_NRA
UG_UNKN
UG_WHITENH
UG_BLACKNH
UG_API
UG_AIANOLD
UG_HISPOLD


TUITIONFEE_IN (in-state tuit)
TUITIONFEE_OUT (out of state tuit)
PCTPELL (% students receiving pell)
PCTFLOAN (% students receiving a student federal loan)
GRAD_DEBT_MDN (The median debt for students who have completed)
FAMINC (avg family income)
MD_FAMINC (median family income)

```{r}
college <- Most_Recent_Cohorts_Institution %>%
  select(
    ADM_RATE,
    UGDS,
    UGDS_MEN,
    UGDS_WOMEN,
    GRADS,
    CCSIZSET,
    UGDS_WHITE,
    UGDS_BLACK,
    UGDS_HISP,
    UGDS_ASIAN,
    UGDS_AIAN,
    UGDS_NHPI,
    UGDS_2MOR,
    UGDS_NRA,
    UGDS_UNKN,
    UGDS_WHITENH,
    UGDS_BLACKNH,
    UGDS_API,
    UGDS_AIANOLD,
    UGDS_HISPOLD,
    UG_NRA,
    UG_UNKN,
    UG_WHITENH,
    UG_BLACKNH,
    UG_API,
    UG_AIANOLD,
    UG_HISPOLD,
    TUITIONFEE_IN,
    TUITIONFEE_OUT,
    PCTPELL,
    PCTFLOAN,
    GRAD_DEBT_MDN,
    FAMINC,
    MD_FAMINC
  )
```

EDA

```{r}
college %>% 
  summarise(
    Average_Tuition_In = mean(TUITIONFEE_IN, na.rm = TRUE),
    Average_Tuition_Out = mean(TUITIONFEE_OUT, na.rm = TRUE),
    Median_Family_Income = median(MD_FAMINC, na.rm = TRUE),
    Average_Pct_Pell = mean(PCTPELL, na.rm = TRUE)
  )
```

```{r}
# Distribution of In-State Tuition Fees
ggplot(college, aes(x = TUITIONFEE_IN)) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7)
# Distribution of Median Family Income
ggplot(college, aes(x = MD_FAMINC)) +
  geom_histogram(bins = 30, fill = "green", alpha = 0.7) +
  labs(title = "Distribution of Median Family Income", x = "Median Family Income", y = "Frequency")
```

```{r}
college %>% 
  mutate(Total_UGDS = UGDS_MEN + UGDS_WOMEN) %>%
  ggplot(aes(x = factor(1), y = UGDS_MEN / Total_UGDS, fill = "Men")) +
  geom_bar(stat = "identity") +
  geom_bar(aes(y = UGDS_WOMEN / Total_UGDS, fill = "Women"), stat = "identity") +
  scale_fill_manual(values = c("Men" = "blue", "Women" = "pink")) +
  labs(x = "", y = "Percentage", title = "Gender Distribution in Undergraduate Enrollment") +
  coord_polar(theta = "y")
```

```{r}
# Checking correlations
numeric_vars <- college %>% 
  select(TUITIONFEE_IN, TUITIONFEE_OUT, MD_FAMINC, PCTPELL, PCTFLOAN)
cor(na.omit(numeric_vars))
```

